---
title: "An Introduction to Rhisat"
author: "Zheng Wei and Wei Zhang"
date: '`r Sys.Date()`'
output:
  html_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{An Introduction to Rhisat}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

The package provides an R wrapper of
[Hisat](http://ccb.jhu.edu/software/hisat/index.shtml) and
[AdapterRemoval](https://github.com/MikkelSchubert/adapterremoval).

Hisat is a fast and sensitive spliced alignment program.
With hierarchical indexing schema, Hisat is much faster >50 
times than TopHat2 with better alignment quality. Hisat performs well 
in the [alignment speed](https://www.nature.com/articles/nmeth.3317/figures/6) 
of the programs in terms of the number of reads processed per second.[1]
AdapterRemoval is a convenient tool for rapid adapter trimming,
identification, and read merging[2]. 

Both of them are implemented with C++. We wrap them
into an R package that provide user friendly interfaces for R users.
Basic functions like "hisat-align" is encapsulated as function "hisat",
"hisat-build" is encapsulated as function "hisat_build". The options in
original binary files are all migrated to Rhisat package. You can also
use function like "hisat_usage" for more help information.

You can preprocess the raw sequencing data by using AadapterRemoval even
if adapter(s) information is missing. Then, Hisat can aligned these
preprocessed reads to the references. 

This package is developed and maintained by members of 
[Xiaowo Wang Lab](http://bioinfo.au.tsinghua.edu.cn/member/xwwang)

MOE Key Laboratory of Bioinformatics and Bioinformatics Division, 

TNLIST / Department of Automation, Tsinghua University

contact:{wei-z14,dqx18}(at)mails.tsinghua.edu.cn

## An Example Workflow by Using Rhisat



### Installation

To install the latest version of Rhisat, you will need to be using the
latest version of R. Rhisat is part of Bioconductor project, 
so you can install Rhisat and its dependencies like this:

```{r install, eval=FALSE}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("Rhisat")
```

### Loading

Just like other R package, 
you need to load Rhisat like this each time before using the package.
```{r loading}
library(Rhisat)
```

### AdapterRemoval

All package functions mentioned in this subsection 
use the binary of AdapterRemoval.

#### Idetitify Adapter

If you know the adapter sequence of reads files, you can skip this step.
Besides,single end data is not support for this function yet so adapter
sequence has to be known .


reads_1 and reads_2 are raw paired-end reads file with fastq format.
adapters is two adapters character vector. 

```{r idad}
td <- tempdir()
reads_1 <- system.file(package="Rhisat", "extdata", "adrm", "reads_1.fq")
reads_2 <- system.file(package="Rhisat", "extdata", "adrm", "reads_2.fq")
(adapters <- 
    identify_adapters(file1=reads_1,file2=reads_2,
                      basename=file.path(td,"reads"),
                      "--threads 3",overwrite=TRUE))
```

#### Remove Adapter

With known adapter sequence, 
remove_adapter function can be call to trim adapters.  

```{r rmad}
(cmdout<-remove_adapters(file1=reads_1,file2=reads_2,adapter1 = adapters[1], 
                adapter2 = adapters[2],
output1=file.path(td,"reads_1.trimmed.fq"),
output2=file.path(td,"reads_2.trimmed.fq"),
basename=file.path(td,"reads.base"),overwrite=TRUE,"--threads 3"))
```


#### Additional Arguments and Version

If you need to set additional arguments like "--threads 3" above, 
you can call function below to print all options available. 
The fixed arguments like file1, file2 and basename etc. are invalid.

```{r adrmusage}
adapterremoval_usage()
```

You can get version information by call:

```{r adrmversion}
adapterremoval_version()
```


### Hisat
All package functions mentioned in this subsection 
use the binary of Hisat. 
Note that Hisat is support 64bit R.

#### Build Hisat Index
Before aligning reads, Hisat index should be build. 
refs is a character vector of fasta reference file paths. 
A prefix of Hisat index should be set to argument bt2Index. 
Then, 6 index files with .bt2 file name extension will 
be created with bt2Index prefix.

Differ from Bowtie2, Hisat build both global and local FM
index. In addition to spliced alignment, HISAT handles reads 
involving indels and supports a paired-end alignment mode. 

```{r hs_bd}
td <- tempdir()
refs <- dir(system.file(package="Rhisat", "extdata", "bt2","refs"),full=TRUE)
(cmdout<-hisat_build(references=refs, 
              bt2Index=file.path(td, "lambda_virus"), "--quiet",
              overwrite=TRUE))
```

#### Additional Arguments of Hisat Build

If you need to set additional arguments like 
"--quiet" above, 
you can call function below to print all options available. 
The fixed arguments references, bt2Index are invalid.

```{r hs_bd_usage}
hisat_build_usage()
```

#### Hisat Alignment

The variable reads_1 and reads_1 are preprocessed reads file paths. 
With hisat index(.bt), 
reads will be mapped to reference by calling hisat. 
The result is saved in a sam file whose path is set to samOutput 

```{r hs_align}
reads_1 <- system.file(package="Rhisat", "extdata", 
                       "bt2", "reads", "reads_1.fastq")
reads_2 <- system.file(package="Rhisat", "extdata", 
                       "bt2", "reads", "reads_2.fastq")
if(file.exists(file.path(td, "lambda_virus.1.bt2"))){
    (cmdout<-hisat(bt2Index = file.path(td, "lambda_virus"),
        samOutput = file.path(td, "result.sam"),
    seq1=reads_1,seq2=reads_2,overwrite=TRUE,"--threads 3"))
    head(readLines(file.path(td, "result.sam")))
}
```

#### Additional Arguments and Version of Hisat Aligner

If you need to set additional arguments like "--threads 3" above, 
you can call function below to print all options available. 
The fixed arguments like bt2Index, samOutput and seq1 etc. are invalid.

```{r hs_usage}
hisat_usage()
```

You can get version information by call:

```{r hs_version}
hisat_version()
```


## Session Infomation

```{r sessioninfo}
sessionInfo()
```

## Acknowledgement

I would strongly appreciate Wei Zheng's work for
package testing and template construction.

## References

[1] Kim, D., Langmead, B. & Salzberg, S. HISAT: a fast 
spliced aligner with low memory requirements. Nat Methods 12, 
357-360 (2015). https://doi.org/10.1038/nmeth.3317

[2] Schubert, Lindgreen, and Orlando (2016). AdapterRemoval v2: rapid
adapter trimming, identification, and read merging. BMC Research Notes,
12;9(1):88.

